### 2020-04-14 paper/main.tex
update template

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2020

% ready for submission
% \usepackage{neurips_2020}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2020}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2020}

% to avoid loading the natbib package, add option nonatbib:
     \usepackage[nonatbib]{neurips_2020}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{Curiosity Driven Reinforcement Learning}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{
  Bob Wei\\
  Department of Computer Science\\
  University of Waterloo\\
  \texttt{q25wei@edu.uwaterloo.ca} \\
  % examples of more authors
  \And
  Akshay Patel \\
  Department of Computer Science\\
  University of Waterloo\\
  \texttt{q25wei@edu.uwaterloo.ca} \\
  \AND
  Samir Alazzam \\
  Department of Computer Science\\
  University of Waterloo\\
  \texttt{q25wei@edu.uwaterloo.ca} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

\section{Submission of papers to NeurIPS 2020}

NeurIPS requires electronic submissions.  The electronic submission site is
\begin{center}
  \url{https://cmt3.research.microsoft.com/NeurIPS2020/}
\end{center}

Please read the instructions below carefully and follow them faithfully.

\subsection{Style}

Papers to be submitted to NeurIPS 2020 must be prepared according to the
instructions presented here. Papers may only be up to eight pages long,
including figures. Additional pages \emph{containing only a section on the broader impact, acknowledgments and/or cited references} are allowed. Papers that exceed eight pages of content will not be reviewed, or in any other way considered for
presentation at the conference.

The margins in 2020 are the same as those in 2007, which allow for $\sim$$15\%$
more words in the paper compared to earlier years.

Authors are required to use the NeurIPS \LaTeX{} style files obtainable at the
NeurIPS website as indicated below. Please make sure you use the current files
and not previous versions. Tweaking the style files may be grounds for
rejection.

\subsection{Retrieval of style files}

The style files for NeurIPS and other conference information are available on
the World Wide Web at
\begin{center}
  \url{http://www.neurips.cc/}
\end{center}
The file \verb+neurips_2020.pdf+ contains these instructions and illustrates the
various formatting requirements your NeurIPS paper must satisfy.

The only supported style file for NeurIPS 2020 is \verb+neurips_2020.sty+,
rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
  Microsoft Word, and RTF are no longer supported!}

The \LaTeX{} style file contains three optional arguments: \verb+final+, which
creates a camera-ready copy, \verb+preprint+, which creates a preprint for
submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
\verb+natbib+ package for you in case of package clash.

\paragraph{Preprint option}
If you wish to post a preprint of your work online, e.g., on arXiv, using the
NeurIPS style, please use the \verb+preprint+ option. This will create a
nonanonymized version of your work with the text ``Preprint. Work in progress.''
in the footer. This version may be distributed as you see fit. Please \textbf{do
  not} use the \verb+final+ option, which should \textbf{only} be used for
papers accepted to NeurIPS.

At submission time, please omit the \verb+final+ and \verb+preprint+
options. This will anonymize your submission and add line numbers to aid
review. Please do \emph{not} refer to these line numbers in your paper as they
will be removed during generation of camera-ready copies.

The file \verb+neurips_2020.tex+ may be used as a ``shell'' for writing your
paper. All you have to do is replace the author, title, abstract, and text of
the paper with your own.

The formatting instructions contained in these style files are summarized in
Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
type with a vertical spacing (leading) of 11~points.  Times New Roman is the
preferred typeface throughout, and will be selected for you by default.
Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
indentation.

The paper title should be 17~point, initial caps/lower case, bold, centered
between two horizontal rules. The top rule should be 4~points thick and the
bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
below the title to rules. All pages should start at 1~inch (6~picas) from the
top of the page.

For the final version, authors' names are set in boldface, and each name is
centered above the corresponding address. The lead author's name is to be listed
first (left-most), and the co-authors' names (if different address) are set to
follow. If there is only one co-author, list both author and co-author side by
side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

All headings should be lower case (except for first word and proper nouns),
flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the heading in
bold, flush left, and inline with the text, with the heading followed by 1\,em
of space.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.  Citations may be
author/year or numeric, as long as you maintain internal consistency.  As to the
format of the references themselves, any style is acceptable as long as it is
used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations appropriate for
use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may add the
following before loading the \verb+neurips_2020+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add the optional
argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{neurips_2020}
\end{verbatim}

As submission is double blind, refer to your own published work in the third
person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
previous work [4].'' If you cite your other papers that are not widely available
(e.g., a journal paper under review), use anonymous author names in the
citation, e.g., an author of the form ``A.\ Anonymous.''

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote, indicate
footnotes with a number\footnote{Sample of the first footnote.} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

All artwork must be neat, clean, and legible. Lines should be dark enough for
purposes of reproduction. The figure number and caption always appear after the
figure. Place one line space before the figure caption and one line space after
the figure. The figure caption should be lower case (except for first word and
proper nouns); figures are numbered consecutively.

You may use color figures.  However, it is best for the figure captions and the
paper body to be legible if the paper is printed in either black/white or in
color.

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table number and
title always appear before the table.  See Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical rules.} We
strongly suggest the use of the \verb+booktabs+ package, which allows for
typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Final instructions}

Do not change any aspects of the formatting parameters in the style files.  In
particular, do not modify the width or length of the rectangle the text should
fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PDF files}

Please prepare submission files with paper size ``US Letter,'' and not, for
example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file must only
contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
  available out-of-the-box on most Linux machines.

\item The IEEE has recommendations for generating PDF files whose fonts are also
  acceptable for NeurIPS. Please see
  \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
  "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
  the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
figure width as a multiple of the line width as in the example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
necessary.

\section*{Broader Impact}

Authors are required to include a statement of the broader impact of their work, including its ethical aspects and future societal consequences.
Authors should discuss both positive and negative outcomes, if any. For instance, authors should discuss a)
who may benefit from this research, b) who may be put at disadvantage from this research, c) what are the consequences of failure of the system, and d) whether the task/method leverages
biases in the data. If authors believe this is not applicable to them, authors can simply state this.

Use unnumbered first level headings for this section, which should go at the end of the paper. {\bf Note that this section does not count towards the eight pages of content that are allowed.}

\begin{ack}
Use unnumbered first level headings for the acknowledgments. All acknowledgments
go at the end of the paper before the list of references. Moreover, you are required to declare
funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2020/PaperInformation/FundingDisclosure}.

Do {\bf not} include this section in the anonymized submission, only in the final paper. You can use the \texttt{ack} environment provided in the style file to autmoatically hide this section in the anonymized submission.
\end{ack}

\section*{References}

References follow the acknowledgments. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references.
{\bf Note that the Reference section does not count towards the eight pages of content that are allowed.}
\medskip

\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.

\end{document}

### 2020-04-15 a2c.py
Update to ICM

@@ -35,7 +35,8 @@ class ActorCritic(nn.Module):
 
         actions_prob = F.softmax(policy_logits, dim=-1)
         actions_distr = torch.distributions.Categorical(actions_prob)
-        next_action = actions_distr.sample()
+        with torch.no_grad():
+            next_action = actions_distr.sample()
 
         return next_action, actions_distr.log_prob(next_action), actions_distr.entropy().mean(), value.squeeze(1)
 
@@ -55,14 +56,17 @@ class A2C(object):
 
         self.num_actions = self.env.action_space.n
         self.actor_critic = ActorCritic(self.num_actions, config).to(device)
-        self.optim = torch.optim.Adam(self.actor_critic.parameters(), lr=config['lr'])
 
         if self.config['use_icm']:
             self.icm_fwd_loss = []
             self.icm_inv_loss = []
             self.curiosity = ICMAgent(self.num_actions, self.config, self.device).to(self.device)
             # lr might have to be diff
-            self.optim_icm = torch.optim.Adam(self.curiosity.parameters(), lr=config['lr'])
+            # self.optim_icm = torch.optim.Adam(self.curiosity.parameters(), lr=0.1)
+            self.cross_entropy_loss = nn.CrossEntropyLoss()
+            self.optim = torch.optim.Adam(list(self.actor_critic.parameters()) + list(self.curiosity.parameters()), lr=config['lr'])
+        else:
+            self.optim = torch.optim.Adam(self.actor_critic.parameters(), lr=config['lr'])
 
     def process_obs(self, obs):
         # 1. reorder dimensions for nn.Conv2d (batch, ch_in, width, height)
@@ -105,23 +109,26 @@ class A2C(object):
             
             if self.config['use_icm']:
                 # ICM stuff here
-                action = torch.zeros((4,self.num_actions), dtype=torch.float32, device=self.device)
-                for i in range(4):
-                    action[i,next_action[i]] = 1
-                ireward, next_pred_state_features, pred_action, next_state_features = self.curiosity.forward(
-                    action, 
-                    self.process_obs(curr_state), 
-                    self.process_obs(obs)
-                    )
-                fwd_model_loss = ((next_pred_state_features-next_state_features)**2).mean()
-                inverse_model_loss = ((pred_action-action)**2).mean()
-                self.icm_fwd_loss += [fwd_model_loss.item()*self.config['fwd_beta']]
[…]

### 2020-04-15 main.py

@@ -4,10 +4,15 @@ import os
 import argparse
 import torch
 import shutil
+import random
+import numpy as np
+
+import tensorflow as tf
 
 from a2c import A2C
 from tensorboardX import SummaryWriter
 
+from stable_baselines.common import set_global_seeds
 from stable_baselines.common.cmd_util import make_atari_env
 from stable_baselines.common.vec_env import VecFrameStack
 
@@ -38,11 +43,25 @@ def get_logger(config):
 
     return logger
 
+def set_seed(seed):
+    set_global_seeds(seed)
+    os.environ['PYTHONHASHSEED'] = str(seed)
+    random.seed(seed)
+    np.random.seed(seed)
+
+    tf.compat.v1.set_random_seed(seed)
+
+    torch.manual_seed(seed)
+    torch.backends.cudnn.deterministic = True
+    torch.backends.cudnn.benchmark = False
+
 def main():
     args = parser.parse_args()
 
     with open(args.config) as f:
         config = yaml.safe_load(f)
+
+    set_seed(42)
 
     writer = None
     # Will ERROR if outdir already exists

### 2020-04-15 a2c.py
Fixes to icm

@@ -203,7 +203,8 @@ class A2C(object):
             #     torch.save(self.actor_critic.state_dict(), os.path.join(self.config['outdir'], 'model_best_ckpt'))
 
             if i % self.config['save_iters'] == 0:
-                torch.save(self.actor_critic.state_dict(), os.path.join(self.config['outdir'], 'model_recent_ckpt'))
+                torch.save(self.actor_critic.state_dict(), os.path.join(self.config['outdir'], 'actor_critic_ckpt_{}'.format(i)))
+                torch.save(self.curiosity.state_dict(), os.path.join(self.config['outdir'], 'icm_ckpt_{}'.format(i)))
 
             # =======================================================================================================
             # LOGGING

### 2020-04-15 main.py

@@ -61,7 +61,7 @@ def main():
     with open(args.config) as f:
         config = yaml.safe_load(f)
 
-    set_seed(42)
+    set_seed(config['seed'])
 
     writer = None
     # Will ERROR if outdir already exists

### 2020-04-15 paper/main.tex
email fix
update paper

@@
 
-\title{Curiosity Driven Reinforcement Learning}
+\title{Motivating Exploration in Reinforcement Learning}
 
@@
   \texttt{q25wei@edu.uwaterloo.ca} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
 }
@@
 
-\section{Submission of papers to NeurIPS 2020}
+\section{Introduction}
 
-NeurIPS requires electronic submissions.  The electronic submission site is
-\begin{center}
-  \url{https://cmt3.research.microsoft.com/NeurIPS2020/}
-\end{center}
+Exploration is something that comes naturally to humans, almost as if there exists an innate sense of curiosity craving to experience something new, something yet to be seen.
+Through observing human behaviour, it is clear that the ability to explore a domain is critical to the process of discovery and learning, regardless of the task being considered.
 
-Please read the instructions below carefully and follow them faithfully.
+And yet, even with the massive computational resources available today, the most impressive artificial learning agents struggle tremendously in balancing explicit exploration of the environment and stable convergence towards a useful, learned policy.
+Without adequate exploration and randomness, the agent can easily become stuck and converge towards a flawed policy early on in learning; the agent's policy reflects high certainty on its understanding of the environment when, in reality, it has yet to explore the majority of the state space.
+On the other hand, Excessive exploration can lead to unstable rewards and updates throughout learning which are detrimental to learning a useful policy.
+To make matters worse, there is a huge breadth of algorithms and hand-crafted techniques used in reinforcement learning, many of them requiring method or task specific hyperparameter tuning.
+This in turn leads to very specialized formulations across the field for encouraging exploration.
 
-\subsection{Style}
+In this work, we explore a general formulation of curiosity that aims to motivate exploration in reinforcement learning (RL) agents while remaining invariant to the algorithm and task in question.
+We explore the effects of including an entropy term (of the policy action probability distribution) and a curiosity module based on that of \textbf{cite}, which provides an intrinsic reward signal.
+Due to the scope of the porject, we focus specifically on the Advantage Actor-Critic (A2C) RL algorithm in the \textit{Pong}, \textit{Seaquest}, and \textit{Breakout} Atari environments.
+These tasks were chosen due to the differences in mechanics and their respective state spaces, specifically the density and magnitude of the extrinsic reward (i.e.
+reward received from environment itself).
 
-Papers to be submitted to NeurIPS 2020 must be prepared according to the
-instructions presented here. Papers may only be up to eight pages long,
-including figures.
-Additional pages \emph{containing only a section on the broader impact, acknowledgments and/or cited references} are allowed.
-Papers that exceed eight pages of content will not be reviewed, or in any other way considered for
-presentation at the conference.
+We compare the performance of the agent in the various environments with and without the mentioned exploration factors. \textbf{Add more on the experimental results once that's finalized}
 
-The margins in 2020 are the same as those in 2007, which allow for $\sim$$15\%$
-more words in the paper compared to earlier years.
+\section{Related Work}
+Someone write this pls
 
-Authors are required to use the NeurIPS \LaTeX{} style files obtainable at the
-NeurIPS website as indicated below. Please make sure you use the current files
-and not previous versions. Tweaking the style files may be grounds for
-rejection.
+\section{Methods}
+In this section, we present our baseline RL framework using the A2C algorithm which is based on the previous works of \textbf{cite}.
+We then describe the formulations of entropy and curiosity based learning factors to encourage environment exploration.
 
-\subsection{Retrieval of style files}
+\subsection{Advantage Actor-Critic (A2C)}
+Our baseline is built around A2C, which is an on-policy learning algorithm.
+The core decision making of the agent stems from the policy network ($\pi$), which is also referred to as the \textit{Actor}.
+The policy network is learned and its weights ($\theta_\pi$) are updated via the standard policy gradient equation, wherein the rewards $r_t$ for a trajectory $t$ are weighted by the negative log likelihood of that trajectory.
+Minimizing this loss $\mathcal{L}_P$ is equivalent to updating $\pi$ such that the probability of high reward trajectories are maximized.
+\begin{equation}\label{eq:policy}
+\mathcal{L}_{P} = \sum_{t} -\log\pi(s_t; \theta_\pi) r_t
+\end{equation}
+A2C also uses the value network ($V$) or \textit{critic} which predicts the accumulated, discounted rewards $R_i$ over the episode timesteps $i$.
+The advantages can be computed as $A_i = R_i - V_i$, where $V_i$ are the predictions from $V$, and are used in place of the actual rewards in the policy update \ref{eq:policy}, reducing the variance of $\mathcal{L}_P$, which is a common downfall of on-policy methods.
+We train the value network $V$ with the update equation \ref{eq:value}.
+\begin{equation}\label{eq:value}
+\mathcal{L}_v = \frac{1}{n}\sum_i (R_i - V_i)^2
+\end{equation}
 
-The style files for NeurIPS and other conference information are available on
-the World Wide Web at
-\begin{center}
-  \url{http://www.neurips.cc/}
-\end{center}
-The file \verb+neurips_2020.pdf+ contains these instructions and illustrates the
-various formatting requirements your NeurIPS paper must satisfy.
+\subsection{Motivating Exploration}
 
-The only supported style file for NeurIPS 2020 is \verb+neurips_2020.sty+,
-rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
-  Microsoft Word, and RTF are no longer supported!}
+\section{Experiments}
 
-The \LaTeX{} style file contains three optional arguments: \verb+final+, which
-creates a camera-ready copy, \verb+preprint+, which creates a preprint for
-submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
-\verb+natbib+ package for you in case of package clash.
+\subsection{Implementation Details}
 
-\paragraph{Preprint option}
-If you wish to post a preprint of your work online, e.g., on arXiv, using the
-NeurIPS style, please use the \verb+preprint+ option. This will create a
-nonanonymized version of your work with the text ``Preprint. Work in progress.''
-in the footer. This version may be distributed as you see fit. Please \textbf{do
-  not} use the \verb+final+ option, which should \textbf{only} be used for
-papers accepted to NeurIPS.
+\subsection{Quantitative Results}
 
-At submission time, please omit the \verb+final+ and \verb+preprint+
-options. This will anonymize your submission and add line numbers to aid
-review. Please do \emph{not} refer to these line numbers in your paper as they
-will be removed during generation of camera-ready copies.
+\subsection{Qualitative Results}
 
-The file \verb+neurips_2020.tex+ may be used as a ``shell'' for writing your
-paper. All you have to do is replace the author, title, abstract, and text of
-the paper with your own.
+\section{Discussion}
 
-The formatting instructions contained in these style files are summarized in
-Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.
-
-\section{General formatting instructions}
 \label{gen_inst}
@@
 
-\section{Headings: first level}
-\label{headings}
+\section{Figures}
 
-All headings should be lower case (except for first word and proper nouns),
-flush left, and bold.
-
-First-level headings should be in 12-point type.
-
-\subsection{Headings: second level}
-
-Second-level headings should be in 10-point type.
-
-\subsubsection{Headings: third level}
-
-Third-level headings should be in 10-point type.
-
-\paragraph{Paragraphs}
-
-There is also a \verb+\paragraph+ command available, which sets the heading in
-bold, flush left, and inline with the text, with the heading followed by 1\,em
-of space.
-
-\section{Citations, figures, tables, references}
-\label{others}
-
-These instructions apply to everyone.
-
-\subsection{Citations within the text}
-
-The \verb+natbib+ package will be loaded for you by default.  Citations may be
-author/year or numeric, as long as you maintain internal consistency.  As to the
-format of the references themselves, any style is acceptable as long as it is
-used consistently.
-
-The documentation for \verb+natbib+ may be found at
-\begin{center}
-  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
-\end{center}
-Of note is the command \verb+\citet+, which produces citations appropriate for
-use in inline text.  For example,
-\begin{verbatim}
-   \citet{hasselmo} investigated\dots
-\end{verbatim}
-produces
-\begin{quote}
-  Hasselmo, et al.\ (1995) investigated\dots
-\end{quote}
-
-If you wish to load the \verb+natbib+ package with options, you may add the
-following before loading the \verb+neurips_2020+ package:
-\begin{verbatim}
-   \PassOptionsToPackage{options}{natbib}
-\end{verbatim}
-
-If \verb+natbib+ clashes with another package you load, you can add the optional
-argument \verb+nonatbib+ when loading the style file:
-\begin{verbatim}
-   \usepackage[nonatbib]{neurips_2020}
-\end{verbatim}
-
-As submission is double blind, refer to your own published work in the third
-person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
-previous work [4].'' If you cite your other papers that are not widely available
-(e.g., a journal paper under review), use anonymous author names in the
-citation, e.g., an author of the form ``A.\ Anonymous.''
-
-\subsection{Footnotes}
-
-Footnotes should be used sparingly.  If you do require a footnote, indicate
-footnotes with a number\footnote{Sample of the first footnote.} in the
-text. Place the footnotes at the bottom of the page on which they appear.
-Precede the footnote with a horizontal rule of 2~inches (12~picas).
-
-Note that footnotes are properly typeset \emph{after} punctuation
-marks.\footnote{As in this example.}
-
-\subsection{Figures}
-
-\begin{figure}
+\begin{figure}[h]
   \centering
@@
 
-\subsection{Tables}
+\section{Tables}
 
-All tables must be centered, neat, clean and legible.  The table number and
-title always appear before the table.  See Table~\ref{sample-table}.
-
-Place one line space before the table title, one line space after the
-table title, and one line space after the table. The table title must
-be lower case (except for first word and proper nouns); tables are
-numbered consecutively.
-
-Note that publication-quality tables \emph{do not contain vertical rules.} We
-strongly suggest the use of the \verb+booktabs+ package, which allows for
-typesetting high-quality, professional tables:
-\begin{center}
-  \url{https://www.ctan.org/pkg/booktabs}
-\end{center}
-This package was used to typeset Table~\ref{sample-table}.
-
-\begin{table}
+\begin{table}[h]
   \caption{Sample table title}
@@
 
-\section{Final instructions}
-
-Do not change any aspects of the formatting parameters in the style files.  In
-particular, do not modify the width or length of the rectangle the text should
-fit into, and do not change font sizes (except perhaps in the
-\textbf{References} section; see below). Please note that pages should be
-numbered.
-
-\section{Preparing PDF files}
-
-Please prepare submission files with paper size ``US Letter,'' and not, for
-example, ``A4.''
-
-Fonts were the main cause of problems in the past years. Your PDF file must only
-contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
-achieve this.
-
-\begin{itemize}
-
-\item You should directly generate PDF files using \verb+pdflatex+.
-
-\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
-  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
-  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
-  available out-of-the-box on most Linux machines.
-
-\item The IEEE has recommendations for generating PDF files whose fonts are also
-  acceptable for NeurIPS. Please see
-  \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}
-
-\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
-  "solid" shapes instead.
-
-\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
-  the equivalent AMS Fonts:
-\begin{verbatim}
-   \usepackage{amsfonts}
-\end{verbatim}
-followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
-for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
-workaround for reals, natural and complex:
-\begin{verbatim}
-   \newcommand{\RR}{I\!\!R} %real numbers
-   \newcommand{\Nat}{I\!\!N} %natural numbers
-   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
-\end{verbatim}
-Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.
-
-\end{itemize}
-
-If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
-you to fix it.
-
-\subsection{Margins in \LaTeX{}}
-
-Most of the margin problems come from figures positioned by hand using
-\verb+\special+ or other commands. We suggest using the command
-\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
-figure width as a multiple of the line width as in the example below:
-\begin{verbatim}
-   \usepackage[pdftex]{graphicx} ...
-   \includegraphics[width=0.8\linewidth]{myfile.pdf}
-\end{verbatim}
-See Section 4.4 in the graphics bundle documentation
-(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})
-
-A number of width problems arise when \LaTeX{} cannot properly hyphenate a
-line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
-necessary.
-
-\section*{Broader Impact}
-
-Authors are required to include a statement of the broader impact of their work, including its ethical aspects and future societal consequences.
-Authors should discuss both positive and negative outcomes, if any. For instance, authors should discuss a)
-who may benefit from this research, b) who may be put at disadvantage from this research, c) what are the consequences of failure of the system, and d) whether the task/method leverages
-biases in the data. If authors believe this is not applicable to them, authors can simply state this.
-
-Use unnumbered first level headings for this section, which should go at the end of the paper. {\bf Note that this section does not count towards the eight pages of content that are allowed.}
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2020/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to autmoatically hide this section in the anonymized submission.
-\end{ack}
-
 \section*{References}
 
-References follow the acknowledgments. Use unnumbered first-level heading for
-the references. Any choice of citation style is acceptable as long as you are
-consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
-when listing the references.
-{\bf Note that the Reference section does not count towards the eight pages of content that are allowed.}
 \medskip

### 2020-04-15 play.py
play script

@@ -13,7 +13,7 @@ from tqdm import tqdm
 from stable_baselines.common.vec_env import VecVideoRecorder, DummyVecEnv
 
 video_folder = 'videos/'
-video_length = 10
+video_length = 10 # change for desired video length
 
 parser = argparse.ArgumentParser(description="Curiosity-driven A2C")
 parser.add_argument('--config', default='configs/main.yaml')
@@ -31,12 +31,12 @@ env = VecVideoRecorder(env, video_folder,
 obs = env.reset()
 n = env.action_space.n
 
-model = ActorCritic(n, config)
 device = torch.device('cuda') if config['use_gpu'] else torch.device('cpu')
+model = ActorCritic(n, config).to(device)
 model.load_state_dict(torch.load('checkpoints\pong_noEnt\model_recent_ckpt', map_location=device))
 model.eval()
 
-for i in tqdm(range(video_length+1)):
+for i in tqdm(range(video_length)):
     #env.render(mode='rgb_array')
     tensor = torch.from_numpy(obs.astype(np.float32).transpose((0, 3, 1, 2))) / 255
     tensor = torch.nn.functional.interpolate(tensor, scale_factor=48/tensor.shape[-1])

### 2020-04-15 paper/main.tex
Finished methods

@@
   University of Waterloo\\
-  \texttt{q25wei@edu.uwaterloo.ca} \\
+  \texttt{akshay.patel@uwaterloo.ca} \\
   \AND
@@
 
-Exploration is something that comes naturally to humans, almost as if there exists an innate sense of curiosity craving to experience something new, something yet to be seen.
+Exploration is something that comes naturally to humans, almost as if there exists an innate sense of curiosity, craving to experience something new, something yet to be seen.
 Through observing human behaviour, it is clear that the ability to explore a domain is critical to the process of discovery and learning, regardless of the task being considered.
@@
 Minimizing this loss $\mathcal{L}_P$ is equivalent to updating $\pi$ such that the probability of high reward trajectories are maximized.
-\begin{equation}\label{eq:policy}
-\mathcal{L}_{P} = \sum_{t} -\log\pi(s_t; \theta_\pi) r_t
+\begin{equation}
+\mathcal{L}_{P} = \sum_{t} -\log\pi(s_t; \theta_\pi) \times r_t\label{eq:policy}
 \end{equation}
@@
 The advantages can be computed as $A_i = R_i - V_i$, where $V_i$ are the predictions from $V$, and are used in place of the actual rewards in the policy update \ref{eq:policy}, reducing the variance of $\mathcal{L}_P$, which is a common downfall of on-policy methods.
-We train the value network $V$ with the update equation \ref{eq:value}.
-\begin{equation}\label{eq:value}
-\mathcal{L}_v = \frac{1}{n}\sum_i (R_i - V_i)^2
+We train the value network $V$ with update equation \ref{eq:value}.
+\begin{equation}
+\mathcal{L}_V = \frac{1}{n}\sum_i (R_i - V_i)^2\label{eq:value}
+\end{equation}
+The base A2C algorithm does not explicitly seek random exploration as it learns to decide on the action at a https://arxiv.org/pdf/1601.06733.pdf state purely from the policy update.
+The most commonly used method to introduce randomness to the agent is through maximizing the entropy of the next-action sampling distribution.
+This distribution is categorical with probabilities defined by the policy network logits.
+\begin{equation}\label{eq:entropy}
+\mathcal{L}_E = \sum_a (-\log \pi_a \times \mu_\pi)
+\end{equation}
+Instead of minizing the loss function in \ref{eq:entropy}, we maximize the value so that the individual probabilities of the sampling distribution $\pi_a$ are as similar as possible; thus, inducing a degree of randomness in the policy network's state to action mapping.
+
+\subsection{Intrinsic Curiosity}
+Entropy encourages the agent to explore but in a purely random manner, which has been shown to produce exceptional results (as below).
+However, we also compare with a more targeted approach by leveraging the Intrinsic Curiosity Module (ICM) introduced in \textbf{cite}, which provides an intrinsic reward as a learning signal.
+ICM consists of two linear networks referred to as the forward and inverse models.
+The forward model $f_{Fwd}$ is trained to predict the next state ($\phi(s_{t+1})$) given the current state ($\phi(s_t)$) and the action ($a_t$) decided by the policy net.
+Note that we use $\phi$ to denote a feature extractor network used to encode the raw states.
+\begin{equation}\label{eq:intrinsic}
+r_i = \sum_t (\phi(s_{t+1}) - f_{Fwd}(\phi(s_t);a_t))^2
+\end{equation}
+\begin{equation}\label{eq:reward}
+r = r_e + \beta r_i
+\end{equation}
+The new intrinsic reward $r_i$ is defined as the mean-squared-error of the predicted next state and the actual next state, which is then combined with the extrinsic reward to obtain the total reward used in the policy update \ref{eq:policy}.
+Thus, our agent learns to maximize the likelihood of following trajectories that maximize the intrinsic reward, which means that the agent is encouraged to enter new, unseen states that $f_{Fwd}$ is unable to predict accurately.
+The forward model itself is trained using the loss below.
+\begin{equation}\label{eq:forward}
+\mathcal{L}_{Fwd} = \frac{1}{n} \sum_t (\phi(s_{t+1}) - f_{Fwd}(\phi(s_t);a_t))^2
+\end{equation}
+Further, the purpose of the inverse model $f_{Inv}$ is to train the feature extractor $\phi$ to encode observed features from the raw state that are most impacted by the agent's actions.
+Specifically, the inverse model takes the current and next states and infers the corresponding action taken: $f_{Inv}(\phi(s_t), \phi(s_{t+1})) = \hat{p}$, where $\hat{p}$ is a vector of probabilities for each possible action.
+This is trained with the following,
+\begin{equation}\label{eq:inverse}
+\mathcal{L}_{Inv} = -\log \Big(\frac{e^{\hat{p}[a_t]}}{\sum_j e^{\hat{p}[j]}}\Big) = -\hat{p}[a_t] + \log\Big(\sum_j e^{\hat{p}[j]}\Big)
+\end{equation}
+The described curiosity model is incorporated into the base A2C model and can be trained jointly:
+\begin{equation}
+\mathcal{L} = \mathcal{L}_P + \lambda_{V}\mathcal{L}_V - \lambda_{E}\mathcal{L}_E + \mathcal{L}_{Fwd} + \mathcal{L}_{Inv}
 \end{equation}
 
-\subsection{Motivating Exploration}
-
 \section{Experiments}
+Below, we provide details on our implementation and setup for training the agent.
+We also include plots of the smoothed, extrinsic rewards from full episodes throughout the training, where we compare the effectiveness of motivating random or targeted exploration.
+Finally, we show some qualitative results across various environments that demonstrate interesting agent behaviour.
 

### 2020-04-16 paper/main.tex
Updated impl details

@@
 \end{equation}
+where we use $\lambda_V=0.5$ and $\lambda_E=0.02$.
 
@@
 \subsection{Implementation Details}
+Our work is implemented in Pytorch and we iterate our agents in the Atari environments provided through OpenAI Gym \textbf{cite}, namely \texttt{PongNoFrameskip-v0}, \texttt{SeaquestNoFrameskip-v0}, and \texttt{BreakoutNoFrameskip-v0} (using consistent random seeds for each of the environments).
+Specifically, we train using 4 parallel environments at a time and with each state consisting of the current frame concatenated with 3 previous frames.
+We use the Adam optimizer \textbf{cite} with a learning rate of 0.0001, a rollout size of 5, and a varying number of update steps (due to the limited scope of the project).
+All of our models are trained synchronously on Nvidia GTX 1080Ti (11GB VRAM) and Tesla V100 (16GB VRAM) gpus with batch size equivalent to the number of parallel environments.
+
+The main actor-critic network consists of a feature extractor, followed by the actor and critic headers, which are simply single linear layers that map the extracted features to the desired next-action logits and predicted discounted rewards respectively.
+The feature extractor consists of four stride-2 convolutional layers followed by a LSTM layer \textbf{cite}, which is needed to leverage past information.
+We note that the LSTM memory is cleared at the end of every episode.
+The encoded features are flattened to $B\times 288$ tensors before being forwarded to the actor-critic.
+The initial state is resized to a $48\times 48$ image before being passed to the feature extractor.
+
+The curiosity module being used consists of a forward and inverse model as well as another feature extractor net (this one being purely convolutional).
+Both forward and inverse models are simple multi-layer perceptrons comprised of two stacked linear layers with 288 hidden neurons.
 
 \subsection{Quantitative Results}
-
-\subsection{Qualitative Results}
-
-\section{Discussion}
-
-\label{gen_inst}
-
-The text must be confined within a rectangle 5.5~inches (33~picas) wide and
-9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
-type with a vertical spacing (leading) of 11~points.  Times New Roman is the
-preferred typeface throughout, and will be selected for you by default.
-Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
-indentation.
-
-The paper title should be 17~point, initial caps/lower case, bold, centered
-between two horizontal rules. The top rule should be 4~points thick and the
-bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
-below the title to rules. All pages should start at 1~inch (6~picas) from the
-top of the page.
-
-For the final version, authors' names are set in boldface, and each name is
-centered above the corresponding address. The lead author's name is to be listed
-first (left-most), and the co-authors' names (if different address) are set to
-follow. If there is only one co-author, list both author and co-author side by
-side.
-
-Please pay special attention to the instructions in Section \ref{others}
-regarding figures, tables, acknowledgments, and references.
-
-\section{Figures}
-
-\begin{figure}[h]
+\begin{figure}[ht]\label{fig:pong-reward}
   \centering
@@
 
-All artwork must be neat, clean, and legible. Lines should be dark enough for
-purposes of reproduction. The figure number and caption always appear after the
-figure. Place one line space before the figure caption and one line space after
-the figure. The figure caption should be lower case (except for first word and
-proper nouns); figures are numbered consecutively.
+\subsection{Qualitative Results}
 
-You may use color figures.  However, it is best for the figure captions and the
-paper body to be legible if the paper is printed in either black/white or in
-color.
-
-\section{Tables}
-
-\begin{table}[h]
-  \caption{Sample table title}
-  \label{sample-table}
-  \centering
-  \begin{tabular}{lll}
-    \toprule
-    \multicolumn{2}{c}{Part}                   \\
-    \cmidrule(r){1-2}
-    Name     & Description     & Size ($\mu$m) \\
-    \midrule
-    Dendrite & Input terminal  & $\sim$100     \\
-    Axon     & Output terminal & $\sim$10      \\
-    Soma     & Cell body       & up to $10^6$  \\
-    \bottomrule
-  \end{tabular}
-\end{table}
+\section{Discussion}
 

### 2020-04-16 paper/main.tex
related work

@@
 \section{Related Work}
-Someone write this pls
+Our artificial agent is a model-free, on-policy reinforcement learning algorithm that leverages Advantage Actor Critic (A2C) methods, entropy-based exploration methods, and curiosity-driven exploration methods.
+Our artificial agent uses combination(s) of these methods to learn how to navigate different OpenAI Gym (cite?) environments to maximize cumulative rewards.
+
+Actor Critic methods have been successfully applied to a variety of reinforcement learning problems, with Advantage Actor Critic methods also being used after their introduction in 2016 through Schulman et al.’s work, High Dimensional Continuous Control Using Generalized Advantage Estimation, done at the University of California, Berkeley.
+Schulman et al.’s work presents the Advantage Actor Critic method that our artificial agent primarily leverages to learn how to maximize cumulative rewards in a variety of environments.
+Schulman et al.
+address two main challenges faced by policy gradient methods in reinforcement learning, namely the large number of samples usually required and the difficulty of gaining consistent improvement with nonstationarity of the data [cite].
+
+Curiosity-driven exploration enhancements to A2C methods have also recently been applied to a set of reinforcement learning algorithms after their introduction in 2017 through Pathak et al.’s work, Curiosity-driven Exploration by Self-supervised Prediction, which was also done at the University of California, Berkeley.
+Pathak et al.’s work presents the curiosity-driven exploration method that our artificial agent uses, namely the Intrinsic Curiosity Module (ICM).
+Pathak et al.’s work addresses a challenge faced in learning when extrinsic rewards for the agent are extremely sparse through curiosity as the error in an agent’s ability to to predict the consequence of its own actions [cite].
+Their proposed ICM notably ignores the aspects of the environment that do not affect the agent, as well as being able to scale to high-dimensional continuous state spaces such as images, and bypasses the difficulties of directly predicting pixels [cite].
+
+Attention based enhancements to deep learning algorithms have also been used to solve reinforcement learning problems more efficiently.
+Reizinger et al.’s work, Attention-Based Curiosity-Driven Exploration In Deep Reinforcement Learning, done at the Budapest University of Technology and Economics, introduces an attention-based enhancement to A2C [cite].
 

### 2020-04-16 a2c.py
Comments and plot code

@@ -61,10 +61,11 @@ class A2C(object):
             self.icm_fwd_loss = []
             self.icm_inv_loss = []
             self.curiosity = ICMAgent(self.num_actions, self.config, self.device).to(self.device)
-            # lr might have to be diff
-            # self.optim_icm = torch.optim.Adam(self.curiosity.parameters(), lr=0.1)
             self.cross_entropy_loss = nn.CrossEntropyLoss()
-            self.optim = torch.optim.Adam(list(self.actor_critic.parameters()) + list(self.curiosity.parameters()), lr=config['lr'])
+            self.optim = torch.optim.Adam(
+                list(self.actor_critic.parameters()) + list(self.curiosity.parameters()), 
+                lr=config['lr'])
+        
         else:
             self.optim = torch.optim.Adam(self.actor_critic.parameters(), lr=config['lr'])
 
@@ -93,6 +94,10 @@ class A2C(object):
         return out
 
     def run_episode(self, obs):
+        """
+        Runs a single episode of length "rollout_steps"
+        """
+        # storage for current episode
         rollout = Storage(self.config['parallel_envs'])
         
         for i in range(self.config['rollout_steps']):
@@ -109,37 +114,32 @@ class A2C(object):
             
             if self.config['use_icm']:
                 # ICM stuff here
-                one_hot_action = torch.zeros((self.config['parallel_envs'], self.num_actions), dtype=torch.float32, device=self.device)#.scatter_(1, next_action, 1)
+
+                # turn next action index into one hot vector
+                one_hot_action = torch.zeros((self.config['parallel_envs'], self.num_actions), dtype=torch.float32, device=self.device)
                 for i in range(self.config['parallel_envs']):
                     one_hot_action[i, next_action[i]] = 1
 
+                # get intrinsic reward and outputs from ICM forward and inverse models
                 ireward, next_pred_state_features, pred_action_logits, next_state_features = self.curiosity(
                     one_hot_action, self.process_obs(curr_state), self.process_obs(obs))
 
-                fwd_model_loss = ((next_pred_state_features - next_state_features)**2).mean()
-                inv_model_loss = self.cross_entropy_loss(pred_action_logits, next_action)
-                # self.icm_fwd_loss += [fwd_model_loss.item()*self.config['fwd_beta']]
-                # self.icm_inv_loss += [inverse_model_loss.item()*(1-self.config['fwd_beta'])]
-                self.icm_fwd_loss.append(fwd_model_loss)
[…]

### 2020-04-16 env_play.py

@@ -1,20 +1,0 @@
-import gym
-import random
-import numpy as np
-
-env = gym.make('SkiingNoFrameskip-v4')
-obs = env.reset()
-n = env.action_space.n
-for i in range(10000):
-    env.render()
-    action = random.randint(0,n-1)
-    obs, rewards, dones, infos = env.step(action)
-    print(rewards)
-if 'episode' in infos.keys():
-    print(infos['episode']['r'])
-
-if dones > 0:
-    print("hello")
-    obs = env.reset()
-else:
-    print("render")

### 2020-04-16 plot.py

@@ -1,0 +1,40 @@
+import matplotlib.pyplot as plt
+import argparse
+import os
+import csv
+import pandas as pd
+
+parser = argparse.ArgumentParser()
+parser.add_argument('--name')
+parser.add_argument('--paths', nargs='+')
+args = parser.parse_args()
+
+# exponential moving average
+def smooth(scalars, weight):
+    last = scalars[0]
+    smoothed = list()
+    for point in scalars:
+        smoothed_val = last * weight + (1 - weight) * point 
+        smoothed.append(smoothed_val)
+        last = smoothed_val
+    return smoothed
+
+
+col_names = ['Step', 'Value']
+data = []
+for path in args.paths:
+    df = pd.read_csv(path, names=col_names, skipinitialspace=True)
+    data.append((df.Step.values[1:].astype(int), df.Value.values[1:].astype(float)))
+
+colours = ['r', 'y', 'c']
+labels = ['A2C w/o Entropy', 'A2C', 'A2C w/ Curiosity']
+i = 0
+for d in data:
+    plt.plot(d[0], d[1], alpha=0.2, color=colours[i])
+    plt.plot(d[0], smooth(d[1], 0.95), color=colours[i], label=labels[i])
+    i = (i+1) % 3
+plt.legend()
+plt.xlabel('Training Update Steps')
+plt.ylabel('Extrinsic Reward per Episode')
+plt.title(args.name)
+plt.show()

### 2020-04-16 a2c.py
fix use_icm

@@ -177,8 +177,11 @@ class A2C(object):
             value_net_loss = (advantages ** 2).mean() * self.config['value_beta']
             entropy = entropies.sum() * self.config['entropy_beta']
 
-            fwd_loss = sum(self.icm_fwd_loss) / len(self.icm_fwd_loss) # mean over list
-            inv_loss = sum(self.icm_inv_loss) / len(self.icm_inv_loss) # mean over list
+            if self.config['use_icm']:
+                fwd_loss = sum(self.icm_fwd_loss) / len(self.icm_fwd_loss) # mean over list
+                inv_loss = sum(self.icm_inv_loss) / len(self.icm_inv_loss) # mean over list
+                self.icm_fwd_loss = []
+                self.icm_inv_loss = []
 
             loss = policy_net_loss + value_net_loss - entropy + (fwd_loss + inv_loss)
 
@@ -188,9 +191,6 @@ class A2C(object):
             # clip gradients for better stability
             nn.utils.clip_grad_norm_(self.actor_critic.parameters(), self.config['max_grad_norm'])
             self.optim.step()
-
-            self.icm_fwd_loss = []
-            self.icm_inv_loss = []
             
             # =======================================================================================================
             # SAVE CHECKPOINTS

### 2020-04-16 plot.py
updates to paper

@@ -27,7 +27,7 @@ for path in args.paths:
     data.append((df.Step.values[1:].astype(int), df.Value.values[1:].astype(float)))
 
 colours = ['r', 'y', 'c']
-labels = ['A2C w/o Entropy', 'A2C', 'A2C w/ Curiosity']
+labels = ['A2C', 'A2C + Entropy', 'A2C + Entropy + Curiosity']
 i = 0
 for d in data:
     plt.plot(d[0], d[1], alpha=0.2, color=colours[i])

### 2020-04-16 paper/main.tex
Comments and plot code

--- 2020-04-16 paper/main.tex
@@
 \usepackage{microtype}      % microtypography
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 \begin{abstract}
-  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
-  both the left- and right-hand margins. Use 10~point type, with a vertical
-  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
-  bold, and in point size 12. Two line spaces precede the abstract. The abstract
-  must be limited to one paragraph.
+Self-motivation and curiosity come second nature to humans, both of which are crucial in quickly adapting to new environments and achieving distant goals.
+Using this as inspiration for artificial agents trained with reinforcement learning, we present an analysis of different techniques used by the current state-of-the-art to encourage exploration in varying environments.
+Specifically, we focus on formulations for entropy and curiosity, which promote exploration via randomness and state-based novelty respectively.
+We show that such methods are invariant to task/environment specifics, such as the reward density, and we also demonstrate that curious learners train faster based on extrinsic rewards and game visualizations.
 \end{abstract}
@@
 In this work, we explore a general formulation of curiosity that aims to motivate exploration in reinforcement learning (RL) agents while remaining invariant to the algorithm and task in question.
-We explore the effects of including an entropy term (of the policy action probability distribution) and a curiosity module based on that of \textbf{cite}, which provides an intrinsic reward signal.
+We explore the effects of including an entropy term (of the policy action probability distribution) and a curiosity module based on that of \cite{curiosity}, which provides an intrinsic reward signal.
 Due to the scope of the porject, we focus specifically on the Advantage Actor-Critic (A2C) RL algorithm in the \textit{Pong}, \textit{Seaquest}, and \textit{Breakout} Atari environments.
@@
 \section{Related Work}
-Someone write this pls
+Our artificial agent is a model-free, on-policy reinforcement learning algorithm that leverages Advantage Actor Critic (A2C) methods, entropy-based exploration methods, and curiosity-driven exploration methods.
+Our artificial agent uses combination(s) of these methods to learn how to navigate different OpenAI Gym \cite{gym} environments to maximize cumulative rewards.
+
+Actor Critic methods have been successfully applied to a variety of reinforcement learning problems, with Advantage Actor Critic methods also being used after their introduction in 2016 through Schulman et al.’s work, High Dimensional Continuous Control Using Generalized Advantage Estimation, done at the University of California, Berkeley.
+Schulman et al.’s work presents the Advantage Actor Critic method that our artificial agent primarily leverages to learn how to maximize cumulative rewards in a variety of environments \cite{a2c}.
+Schulman et al.
+address two main challenges faced by policy gradient methods in reinforcement learning, namely the large number of samples usually required and the difficulty of gaining consistent improvement with nonstationarity of the data \cite{a2c}.
+
+Curiosity-driven exploration enhancements to A2C methods have also recently been applied to a set of reinforcement learning algorithms after their introduction in 2017 through Pathak et al.’s work, Curiosity-driven Exploration by Self-supervised Prediction, which was also done at the University of California, Berkeley.
+Pathak et al.’s work presents the curiosity-driven exploration method that our artificial agent uses, namely the Intrinsic Curiosity Module (ICM).
+Pathak et al.’s work addresses a challenge faced in learning when extrinsic rewards for the agent are extremely sparse through curiosity as the error in an agent’s ability to to predict the consequence of its own actions \cite{curiosity}.
+Their proposed ICM notably ignores the aspects of the environment that do not affect the agent, as well as being able to scale to high-dimensional continuous state spaces such as images, and bypasses the difficulties of directly predicting pixels \cite{curiosity}.
+
+Attention based enhancements to deep learning algorithms have also been used to solve reinforcement learning problems more efficiently.
+Reizinger et al.’s work, Attention-Based Curiosity-Driven Exploration In Deep Reinforcement Learning, done at the Budapest University of Technology and Economics, introduces an attention-based enhancement to A2C \cite{budapest}.
 
 \section{Methods}
-In this section, we present our baseline RL framework using the A2C algorithm which is based on the previous works of \textbf{cite}.
+In this section, we present our baseline RL framework using the A2C algorithm which is based on the previous works of \cite{a2c}.
 We then describe the formulations of entropy and curiosity based learning factors to encourage environment exploration.
@@
 \subsection{Intrinsic Curiosity}
-Entropy encourages the agent to explore but in a purely random manner, which has been shown to produce exceptional results (as below).
-However, we also compare with a more targeted approach by leveraging the Intrinsic Curiosity Module (ICM) introduced in \textbf{cite}, which provides an intrinsic reward as a learning signal.
+Entropy encourages the agent to explore but in a purely random manner, which has been shown to produce exceptional results (refer to experiments).
+However, we also compare with a more targeted approach by leveraging the Intrinsic Curiosity Module (ICM) introduced in \cite{curiosity}, which provides an intrinsic reward as a learning signal.
 ICM consists of two linear networks referred to as the forward and inverse models.
@@
 \subsection{Implementation Details}
-Our work is implemented in Pytorch and we iterate our agents in the Atari environments provided through OpenAI Gym \textbf{cite}, namely \texttt{PongNoFrameskip-v0}, \texttt{SeaquestNoFrameskip-v0}, and \texttt{BreakoutNoFrameskip-v0} (using consistent random seeds for each of the environments).
+Our work is implemented in Pytorch and we iterate our agents in the Atari environments provided through OpenAI Gym \cite{ppo, gym}, namely \texttt{PongNoFrameskip-v0}, \texttt{SeaquestNoFrameskip-v0}, and \texttt{BreakoutNoFrameskip-v0} (using consistent random seeds for each of the environments).
 Specifically, we train using 4 parallel environments at a time and with each state consisting of the current frame concatenated with 3 previous frames.
-We use the Adam optimizer \textbf{cite} with a learning rate of 0.0001, a rollout size of 5, and a varying number of update steps (due to the limited scope of the project).
+We use the Adam optimizer \cite{adam} with a learning rate of 0.0001, a rollout size of 5, and a varying number of update steps (due to the limited scope of the project).
 All of our models are trained synchronously on Nvidia GTX 1080Ti (11GB VRAM) and Tesla V100 (16GB VRAM) gpus with batch size equivalent to the number of parallel environments.
@@
 The main actor-critic network consists of a feature extractor, followed by the actor and critic headers, which are simply single linear layers that map the extracted features to the desired next-action logits and predicted discounted rewards respectively.
-The feature extractor consists of four stride-2 convolutional layers followed by a LSTM layer \textbf{cite}, which is needed to leverage past information.
+The feature extractor consists of four stride-2 convolutional layers followed by a LSTM layer \cite{lstm}, which is needed to leverage past information.
 We note that the LSTM memory is cleared at the end of every episode.
@@
 \subsection{Quantitative Results}
+Here, we show our training results in the three environments: \texttt{Pong}, \texttt{Seaquest}, and \texttt{Breakout}.
+Specifically, we plot the extrinsic reward (objective measure of agent performance) as training progresses for three different variants of A2C:
+\begin{small}
+\begin{itemize}
+  \item \textbf{A2C}: baseline A2C algorithm without entropy term
+  \item \textbf{A2C + Entropy}: A2C with entropy term to encourage randomized exploration
+  \item \textbf{A2C + Entropy + Curiosity}: A2C with ICM and entropy to encourage exploration of novel states
+\end{itemize}
+\end{small}
+
+\subsubsection{Sparse Extrinsic Reward}
 \begin{figure}[ht]\label{fig:pong-reward}
   \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+  \includegraphics[width=0.6\textwidth]{figures/pong_rewards}
   \caption{Sample figure caption.}
+\end{figure}
+
+\newpage
+\subsubsection{Dense Extrinsic Reward}
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=1.\linewidth]{figures/pong_rewards}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=1.\linewidth]{figures/pong_rewards}
+\end{subfigure}
+\caption{A figure with two subfigures}
+\label{fig:seaquest-breakout-reward}
 \end{figure}
@@
 
-\section*{References}
-
-\medskip
-
-\small
-
-[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
-connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
-(eds.), {\it Advances in Neural Information Processing Systems 7},
-pp.\ 609--616. Cambridge, MA: MIT Press.
-
-[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
-  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
-TELOS/Springer--Verlag.
-
-[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
-recall at excitatory recurrent synapses and cholinergic modulation in rat
-hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
+\bibliography{main}
+\bibliographystyle{ieeetr}
 

### 2020-04-16 paper/main.tex
Update quantitative results

@@
   University of Waterloo\\
-  \texttt{q25wei@edu.uwaterloo.ca} \\
+  \texttt{q25wei@uwaterloo.ca} \\
   % examples of more authors
@@
 A2C also uses the value network ($V$) or \textit{critic} which predicts the accumulated, discounted rewards $R_i$ over the episode timesteps $i$.
-The advantages can be computed as $A_i = R_i - V_i$, where $V_i$ are the predictions from $V$, and are used in place of the actual rewards in the policy update \ref{eq:policy}, reducing the variance of $\mathcal{L}_P$, which is a common downfall of on-policy methods.
-We train the value network $V$ with update equation \ref{eq:value}.
+The advantages can be computed as $A_i = R_i - V_i$, where $V_i$ are the predictions from $V$, and are used in place of the actual rewards in the policy update Eq.
+\ref{eq:policy}, reducing the variance of $\mathcal{L}_P$, which is a common downfall of on-policy methods.
+We train the value network $V$ with update Eq.
+\ref{eq:value}.
 \begin{equation}
@@
 \end{equation}
-Instead of minizing the loss function in \ref{eq:entropy}, we maximize the value so that the individual probabilities of the sampling distribution $\pi_a$ are as similar as possible; thus, inducing a degree of randomness in the policy network's state to action mapping.
+Instead of minizing the loss function in Eq.
+\ref{eq:entropy}, we maximize the value so that the individual probabilities of the sampling distribution $\pi_a$ are as similar as possible; thus, inducing a degree of randomness in the policy network's state to action mapping.
 
@@
 \end{equation}
-The new intrinsic reward $r_i$ is defined as the mean-squared-error of the predicted next state and the actual next state, which is then combined with the extrinsic reward to obtain the total reward used in the policy update \ref{eq:policy}.
+The new intrinsic reward $r_i$ is defined as the mean-squared-error of the predicted next state and the actual next state, which is then combined with the extrinsic reward to obtain the total reward used in the policy update Eq.
+\ref{eq:policy}.
 Thus, our agent learns to maximize the likelihood of following trajectories that maximize the intrinsic reward, which means that the agent is encouraged to enter new, unseen states that $f_{Fwd}$ is unable to predict accurately.
@@
 \end{small}
+Note that in the interest of time, we were unable to perform hyper-parameter tuning for any of our baselines and so the results here are not comparable to other works with finely tuned agents.
+We emphasize the relative performance of the three A2C variants across the three environments, and so we also use the same seed across the board.
 
 \subsubsection{Sparse Extrinsic Reward}
-\begin{figure}[ht]\label{fig:pong-reward}
+\begin{figure}
   \centering
   \includegraphics[width=0.6\textwidth]{figures/pong_rewards}
-  \caption{Sample figure caption.}
+  \caption{We compare the three variants of A2C here, with the curious learner (cyan) vastly outperforming the baseline A2C (red) and the random entropy-based learner (yellow).
+The extrinsic reward is total cumulative reward over a full episode.}
+  \label{fig:pong-reward}
 \end{figure}
 
-\newpage
+Relative to the other environments we tested in, \texttt{Pong} provides a much sparser extrinsic reward signal as a positive reward is provided only when the agent is able to score a point, which is a large hurdle during early training.
+So, in the beginning, much is dependent on the negative reward signal provided only each time the agent allows a point.
+Most environment steps, in fact, provide a zero reward and so the addition of an intrinsic reward (as with ICM) will become the dominant learning signal, pushing the agent to explore unfamiliar states.
+
+As shown in Fig.
+\ref{fig:pong-reward}, the curiosity-driven learner trains much faster than the entropy-based learner, which can be explained by the difference in the style of exploration.
+ICM provides a direct learning signal to the model that encourages the policy to map to actions that are more likely to enter novel, unseen states.
+In contrast, the entropy term aims to distribute the output probabilities over the next-actions in the policy output space as evenly as possible, preventing the agent from becoming stuck but does so by encouraging the actor to select random actions at times.
+This difference perhaps explains the higher potential of the curious model, in that it is able to reach higher cumulative rewards that are highly improbable for the entropy-based model.
+
+Note that the cumulative reward in \texttt{Pong} is the the agent's score minus the cpu's score, and the episode ends when either reaches a score of $21$.
+Thus, the only possible values for cumulative reward are integers between $-21$ and $21$.
+However, a score of $21$ means that the agent did not allow a single point, and so higher rewards are progressively more and more difficult to reach.
+This explains the plateauing of model performance at higher rewards, and perhaps also the entropy-based learner's sudden drop near the end of training.
+
+Looking at the baseline A2C model (red in Fig.
+\ref{fig:pong-reward}), performance is poor due to the fact that the agent has no exploration factor; thus, it tends to get stuck in states where the policy action is to remain in a similar state.
+This can be seen in the portions of the learning curve where performance collapses and remains stuck for some time.
+
 \subsubsection{Dense Extrinsic Reward}
@@
 \end{subfigure}
-\caption{A figure with two subfigures}
+\caption{Here we compare the three A2C variants in the \texttt{Breakout} (left) and \texttt{Seaquest} (right) environments. Rewards are again cumulative over an entire episode.}
 \label{fig:seaquest-breakout-reward}
 \end{figure}
+
+We refer to \texttt{Breakout} and \texttt{Seaquest} as our dense reward environments.
+Specifically, \texttt{Seaquest} is most dense in that positive rewards are given for each \"kill\", which occurs very frequently due to the sheer number of enemies and the fact that the player's submarine is always firing projectiles at high velocity.
+In fact, positive rewards are already obtained in the early steps of training (as can be seen in Fig.
+\ref{fig:seaquest-breakout-reward}).
+\texttt{Breakout} provides a slightly sparser reward but is still denser than that of \texttt{Pong}; rewards are given for each brick broken which occurs every time the agent is able to return the ball.
+This means that in early training, when the agent is unable to consistently return the ball, there is no reward signal and so training performance is heavily dependent on an exploration factor.
+
+We note that curiosity related hyper-parameters were not tuned at all for these environments and so the ICM performance remains an area of improvement, particularly in the variance of the intrinsic rewards.
+This may be the reasoning for the poor performance of the curious learner compared to the entropy-based learner and we hope to address this in future work.
+It can also be said that the benefits of an intrinsic curiosity-driven reward in environments with dense extrinsic rewards are not as substantial as in sparse reward environments, where the agent often depends on the intrinsic reward as the only learning signal.
+
+Further, the baseline A2C models perform very poorly in these environments, as they again tend to get stuck in certain states (this behaviour can be seen later in the qualitative results).
+This leads to sudden mode collapse as in the \texttt{Seaquest} environment and an inability to reach further than a score of 10 in \texttt{Breakout} (sharp plateau in reward learning curve).
+However, relative to the performance in the sparse \texttt{Pong} environment, the baseline A2C model seems to fare better in these dense reward environments thanks to the consistent learning signal.
 

### 2020-04-16 paper/main.tex
Update main.tex

@@
 \subsection{Qualitative Results}
+Qualitative Results
+
+For each trained environment, two agents were trained to investigate the effects of exploration.
+One artificial agent would be trained without exploration and the other would be trained with exploration.
+Videos of the agents playing the game in the environment were generated, and through these videos, qualitative observations are made with regards to the effects of exploration, and any interesting behaviours that are seen in the agents that leverage exploration.
+
+Qualitative Effects of Exploration
+
+Pong
+
+With regards to Pong, the environment with the sparsest rewards from the three that are investigated in this paper, both agents without and with exploration respectively still make use of movement of the paddle to maximize cumulative reward as can be seen in the figures above.
+However, the agent with exploration noticeably has a much higher rate of movement compared to the agent without exploration, which could lead to more states of the environment being explored.
+The agent with exploration often makes many movements that are much faster than the agent without exploration, as can be observed in the supplementary videos.
+
+Breakout
+
+With regards to Breakout, the environment out of the three investigated in this paper with a higher reward density than Pong but a lower reward density than Seaquest, there is a very noticeable difference in the behaviour of the two agents.
+The agent without exploration tends to move to the rightmost corner, and simply stay there for most of the game’s duration as can be seen in the figure above.
+The agent with exploration makes movements all across the screen in order to maximize cumulative reward, and intuitively, explores its environment much more than the agent without exploration.
+This is a very distinct difference in behaviour that can be clearly observed, especially in the supplementary videos provided.
+
+Seaquest
+
+With regards to Seaquest, the environment with the highest reward density, there is also a very noticeable difference in the behaviour of the agents.
+This difference between the two agents is the same as the difference noticed in the agents of Breakout.
+The agent without exploration in Seaquest does not explore its environment at all, and instead opts to stay above sea level at the top left corner.
+This behaviour is quite unique because it completely avoids any negative rewards in the environment caused by the fish attacking the agent or due to oxygen loss.
+On the other hand, the agent with exploration makes movements to fully explore the environment to maximize reward while taking the risk of negative rewards as well.
+
+The differences in behaviour noticed in the agents with and without exploration of both Breakout and Seaquest could perhaps be explained by the lack of maximizing the entropy of the next-action sampling distribution in the agent without exploration.
+Without this maximization of entropy of the next-action sampling distribution, there is no degree of randomness in the policy network’s state to action mapping.
+In turn, the lack of randomness may be causing a collapse in the explorative behaviour due to some, perhaps static, actions having a very high probability defined by the policy network logits.
+
+Interesting Behaviours
+
+Along with the differences observed in the agents with and without exploration, there are also interesting behaviours observed in the agents with exploration with regards to how they interact with their respective environments.
+These can be seen in the supplementary videos provided.
+
+In Pong, it is observed that the agent waits for the ball for a bit, and then rapidly moves towards the ball to hit it with the edge of the paddle in order to launch the ball quickly off of the paddle with a high velocity.
+Perhaps this is a strategy used by the agent to cause a rapid change in velocity that causes the opponent to be unable to catch it in time due to slower reaction speeds.
+
+In Breakout, it is observed that the agent tends to hit the ball from the rightmost corner if it can.
+This might be a strategy used to break apart the bricks at the top left corner in order to push the ball on top of the bricks, and consequently destroy many more bricks from above in order to accumulate much higher rewards.
+
+In Seaquest, it is observed that the agent takes into consideration the oxygen bar, which is a purely visual indicator.
+The agent decides the next actions based on their oxygen remaining as well, opting to return to the surface of the water to regain oxygen in order to further extend the opportunity to accumulate more rewards.
 

### 2020-04-16 paper/main.tex
format figures

@@
 \subsection{Implementation Details}
-Our work is implemented in Pytorch and we iterate our agents in the Atari environments provided through OpenAI Gym \cite{ppo, gym}, namely \texttt{PongNoFrameskip-v0}, \texttt{SeaquestNoFrameskip-v0}, and \texttt{BreakoutNoFrameskip-v0} (using consistent random seeds for each of the environments).
+Our work is implemented in Pytorch and we iterate our agents in the Atari environments provided through OpenAI Gym \cite{ppo, gym}, namely \texttt{PongNoFrameskip-v0}, \texttt{SeaquestNoFrameskip-v0}, and \texttt{BreakoutNoFrameskip-v0}, using consistent random seeds for each of the environments.
 Specifically, we train using 4 parallel environments at a time and with each state consisting of the current frame concatenated with 3 previous frames.
@@
 \subsubsection{Dense Extrinsic Reward}
-\begin{figure}[ht]
+\begin{figure}
 \centering
@@
 \subsection{Qualitative Results}
-Qualitative Results
-
-For each trained environment, two agents were trained to investigate the effects of exploration.
+For each trained environment, two agents were considered to investigate the effects of exploration.
 One artificial agent would be trained without exploration and the other would be trained with exploration.
@@
 
-Qualitative Effects of Exploration
+\subsubsection{Qualitative Effects of Exploration}
 
-Pong
-
-With regards to Pong, the environment with the sparsest rewards from the three that are investigated in this paper, both agents without and with exploration respectively still make use of movement of the paddle to maximize cumulative reward as can be seen in the figures above.
+\subsubsubsection{\textbf{Pong}}
+\begin{figure}
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.7\linewidth]{figures/pong}
+  \caption{With Exploration}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.7\linewidth]{figures/pong_no_exp}
+  \caption{Without Exploration}
+\end{subfigure}
+\caption{Pong environment visualizations}
+\label{fig:pong}
+\end{figure}
+With regards to Pong, the environment with the sparsest rewards from the three that are investigated in this paper, both agents without and with exploration respectively still make use of movement of the paddle to maximize cumulative reward as can be seen in Fig \ref{fig:pong}.
 However, the agent with exploration noticeably has a much higher rate of movement compared to the agent without exploration, which could lead to more states of the environment being explored.
@@
 
-Breakout
-
+\subsubsubsection{\textbf{Breakout}}
+\begin{figure}
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.7\linewidth]{figures/breakout}
+  \caption{With Exploration}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.7\linewidth]{figures/breakout_no_exp}
+  \caption{Without Exploration}
+\end{subfigure}
+\caption{Breakout environment visualizations}
+\label{fig:breakout}
+\end{figure}
 With regards to Breakout, the environment out of the three investigated in this paper with a higher reward density than Pong but a lower reward density than Seaquest, there is a very noticeable difference in the behaviour of the two agents.
-The agent without exploration tends to move to the rightmost corner, and simply stay there for most of the game’s duration as can be seen in the figure above.
+The agent without exploration tends to move to the rightmost corner, and simply stay there for most of the game’s duration as can be seen in Fig.
+\ref{fig:breakout}.
 The agent with exploration makes movements all across the screen in order to maximize cumulative reward, and intuitively, explores its environment much more than the agent without exploration.
@@
 
-Seaquest
-
-With regards to Seaquest, the environment with the highest reward density, there is also a very noticeable difference in the behaviour of the agents.
+\subsubsubsection{\textbf{Seaquest}}
+\begin{figure}
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.7\linewidth]{figures/seaquest}
+  \caption{With Exploration}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.7\linewidth]{figures/seaquest_no_exp}
+  \caption{Without Exploration}
+\end{subfigure}
+\caption{Seaquest environment visualizations}
+\label{fig:seaquest}
+\end{figure}
+With regards to Seaquest, the environment with the highest reward density, there is also a very noticeable difference in the behaviour of the agents as seen in Fig.
+\ref{fig:seaquest}.
 This difference between the two agents is the same as the difference noticed in the agents of Breakout.
@@
 
-Interesting Behaviours
+\subsubsection{Interesting Behaviours}
 

### 2020-04-16 paper/main.tex
fixed e

@@
 Without adequate exploration and randomness, the agent can easily become stuck and converge towards a flawed policy early on in learning; the agent's policy reflects high certainty on its understanding of the environment when, in reality, it has yet to explore the majority of the state space.
-On the other hand, Excessive exploration can lead to unstable rewards and updates throughout learning which are detrimental to learning a useful policy.
+On the other hand, excessive exploration can lead to unstable rewards and updates throughout learning which are detrimental to learning a useful policy.
 To make matters worse, there is a huge breadth of algorithms and hand-crafted techniques used in reinforcement learning, many of them requiring method or task specific hyperparameter tuning.

### 2020-04-16 paper/main.tex
discussion

@@
 \end{figure}
-With regards to Pong, the environment with the sparsest rewards from the three that are investigated in this paper, both agents without and with exploration respectively still make use of movement of the paddle to maximize cumulative reward as can be seen in Fig \ref{fig:pong}.
+With regards to \texttt{Pong}, the environment with the sparsest rewards from the three that are investigated in this paper, both agents without and with exploration respectively still make use of movement of the paddle to maximize cumulative reward as can be seen in Fig \ref{fig:pong}.
 However, the agent with exploration noticeably has a much higher rate of movement compared to the agent without exploration, which could lead to more states of the environment being explored.
@@
 \end{figure}
-With regards to Breakout, the environment out of the three investigated in this paper with a higher reward density than Pong but a lower reward density than Seaquest, there is a very noticeable difference in the behaviour of the two agents.
+With regards to \texttt{Breakout}, the environment out of the three investigated in this paper with a higher reward density than \texttt{Pong} but a lower reward density than \texttt{Seaquest}, there is a very noticeable difference in the behaviour of the two agents.
 The agent without exploration tends to move to the rightmost corner, and simply stay there for most of the game’s duration as can be seen in Fig.
@@
 \end{figure}
-With regards to Seaquest, the environment with the highest reward density, there is also a very noticeable difference in the behaviour of the agents as seen in Fig.
+With regards to \texttt{Seaquest}, the environment with the highest reward density, there is also a very noticeable difference in the behaviour of the agents as seen in Fig.
 \ref{fig:seaquest}.
-This difference between the two agents is the same as the difference noticed in the agents of Breakout.
+This difference between the two agents is the same as the difference noticed in the agents of \texttt{Breakout}.
 The agent without exploration in Seaquest does not explore its environment at all, and instead opts to stay above sea level at the top left corner.
@@
 
-The differences in behaviour noticed in the agents with and without exploration of both Breakout and Seaquest could perhaps be explained by the lack of maximizing the entropy of the next-action sampling distribution in the agent without exploration.
+The differences in behaviour noticed in the agents with and without exploration of both \texttt{Breakout} and \texttt{Seaquest} could perhaps be explained by the lack of maximizing the entropy of the next-action sampling distribution in the agent without exploration.
 Without this maximization of entropy of the next-action sampling distribution, there is no degree of randomness in the policy network’s state to action mapping.
@@
 
-In Pong, it is observed that the agent waits for the ball for a bit, and then rapidly moves towards the ball to hit it with the edge of the paddle in order to launch the ball quickly off of the paddle with a high velocity.
+In \texttt{Pong}, it is observed that the agent waits for the ball for a bit, and then rapidly moves towards the ball to hit it with the edge of the paddle in order to launch the ball quickly off of the paddle with a high velocity.
 Perhaps this is a strategy used by the agent to cause a rapid change in velocity that causes the opponent to be unable to catch it in time due to slower reaction speeds.
 
-In Breakout, it is observed that the agent tends to hit the ball from the rightmost corner if it can.
+In \texttt{Breakout}, it is observed that the agent tends to hit the ball from the rightmost corner if it can.
 This might be a strategy used to break apart the bricks at the top left corner in order to push the ball on top of the bricks, and consequently destroy many more bricks from above in order to accumulate much higher rewards.
 
-In Seaquest, it is observed that the agent takes into consideration the oxygen bar, which is a purely visual indicator.
+In \texttt{Seaquest}, it is observed that the agent takes into consideration the oxygen bar, which is a purely visual indicator.
 The agent decides the next actions based on their oxygen remaining as well, opting to return to the surface of the water to regain oxygen in order to further extend the opportunity to accumulate more rewards.
@@
 \section{Discussion}
+Even with the massive computational resources available today, artificial agents face difficulty in balancing explicit exploration of the environment and stable convergence towards a useful, learned policy.
+In this work, we explored a general formulation of curiosity that aims to motivate the exploration in reinforcement learning agents while remaining invariant to the algorithm and problem in question.
+This general formulation of curiosity was applied to the A2C RL algorithm in various environments with different extrinsic reward densities to examine the effects of an entropy term (of the policy action probability distribution) and a curiosity module providing an intrinsic reward signal.
+
+As discussed in the results section above, the curiosity-driven (ICM) agents in \texttt{Pong} learn and train much faster than the entropy-based agents, which was explained by the difference in the style of exploration.
+Moreover, the baseline A2C model with no exploration performed poorly due to the fact that the agent was not exploring at all.
+With regards to \texttt{Breakout} and \texttt{Seaquest}, particularly dense reward environments, the ICM performance remains an area of improvement through hyper-parameter tuning.
+The baseline A2C models perform poorly in these environments as discussed previously, but they perform better in these environments relative to sparse reward environments like \texttt{Pong}.
+Our results show that exploration factors applied to baseline algorithms can improve model performance and decrease training time through intrinsic reward signals.
+In particular, exploration is crucial across sparse reward environments, which we wanted to verify and which is backed by our results.
+
+The limitations we faced in our work were mostly related to time available to complete this work.
+With more time, we would have liked to train all of  our agents for longer periods of time until convergence.
+Moreover, we would prefer to have done more experiments with more time to observe the quantitative effects of hyperparameter tuning for specific environments such as \texttt{Breakout} and \texttt{Seaquest} and agents with intrinsic curiosity (ICM) since we were unable to do so in the time we had.
+
+In the future, we would like to diversify the environments investigated in this paper.
+\texttt{Pong}, \texttt{Breakout}, and \texttt{Seaquest} are all Atari games where the states and frames of the game do not change drastically from one state to another.
+This could perhaps be a factor affecting the ICM due to the fact that novelty in states for curiosity is defined by the change in its features.
+Through more experiments in a diversified set of environments including environments with more distinct states, additional insights about the effect of exploration and curiosity might be revealed.
 

### 2020-04-16 paper/main.tex

@@
   University of Waterloo\\
-  \texttt{q25wei@edu.uwaterloo.ca} \\
+  \texttt{salazzam@uwaterloo.ca} \\
 }
@@
 Exploration is something that comes naturally to humans, almost as if there exists an innate sense of curiosity, craving to experience something new, something yet to be seen.
-Through observing human behaviour, it is clear that the ability to explore a domain is critical to the process of discovery and learning, regardless of the task being considered.
+Jean Piaget, a Swiss developmental psychologist, asserted that it is in human nature to be curious, inquisitive, and interested in novelty as they develop and acquire knowledge about their environment \cite{piaget}.Through observing human behaviour, it is clear that the ability to explore a domain is critical to the process of discovery and learning, regardless of the task being considered.
 
@@
 We refer to \texttt{Breakout} and \texttt{Seaquest} as our dense reward environments.
-Specifically, \texttt{Seaquest} is most dense in that positive rewards are given for each \"kill\", which occurs very frequently due to the sheer number of enemies and the fact that the player's submarine is always firing projectiles at high velocity.
+Specifically, \texttt{Seaquest} is most dense in that positive rewards are given for each ``kill'', which occurs very frequently due to the sheer number of enemies and the fact that the player's submarine is always firing projectiles at high velocity.
 In fact, positive rewards are already obtained in the early steps of training (as can be seen in Fig.

### 2020-04-16 plot.py
update figures

@@ -14,7 +14,7 @@ def smooth(scalars, weight):
     last = scalars[0]
     smoothed = list()
     for point in scalars:
-        smoothed_val = last * weight + (1 - weight) * point 
+        smoothed_val = last * weight + (1 - weight) * point
         smoothed.append(smoothed_val)
         last = smoothed_val
     return smoothed
@@ -24,8 +24,10 @@ col_names = ['Step', 'Value']
 data = []
 for path in args.paths:
     df = pd.read_csv(path, names=col_names, skipinitialspace=True)
-    data.append((df.Step.values[1:].astype(int), df.Value.values[1:].astype(float)))
+    data.append([df.Step.values[1:].astype(int), df.Value.values[1:].astype(float)])
 
+# data[-1][1] = data[-1][1][:-45]
+# data[-1][0] = data[-1][0][:-45]
 colours = ['r', 'y', 'c']
 labels = ['A2C', 'A2C + Entropy', 'A2C + Entropy + Curiosity']
 i = 0
@@ -33,6 +35,8 @@ for d in data:
     plt.plot(d[0], d[1], alpha=0.2, color=colours[i])
     plt.plot(d[0], smooth(d[1], 0.95), color=colours[i], label=labels[i])
     i = (i+1) % 3
+
+# plt.ylim(0, 1800)
 plt.legend()
 plt.xlabel('Training Update Steps')
 plt.ylabel('Extrinsic Reward per Episode')

### 2020-04-17 paper/main.tex
Update main.tex
update figures

@@
 reward received from environment itself).
+Moreover, our general formulation of curiosity through an entropy term and a curiosity module have not been applied to these specific environments previously.
 
-We compare the performance of the agent in the various environments with and without the mentioned exploration factors. \textbf{Add more on the experimental results once that's finalized}
+We compare the performance of the agent in the various environments with and without the mentioned exploration factors.
+We expected to see better performance in sparse environments with the ICM module, but with regards to denser environments, we expected to see a mitigated effect from the ICM module compared to the sparse environments.
 
@@
   \centering
-  \includegraphics[width=1.\linewidth]{figures/pong_rewards}
+  \includegraphics[width=1.\linewidth]{figures/breakout_rewards}
 \end{subfigure}%
@@
   \centering
-  \includegraphics[width=1.\linewidth]{figures/pong_rewards}
+  \includegraphics[width=1.\linewidth]{figures/seaquest_rewards}
 \end{subfigure}

### 2020-04-17 paper/main.tex
Final sweep of report

@@
 Exploration is something that comes naturally to humans, almost as if there exists an innate sense of curiosity, craving to experience something new, something yet to be seen.
-Jean Piaget, a Swiss developmental psychologist, asserted that it is in human nature to be curious, inquisitive, and interested in novelty as they develop and acquire knowledge about their environment \cite{piaget}.Through observing human behaviour, it is clear that the ability to explore a domain is critical to the process of discovery and learning, regardless of the task being considered.
+Jean Piaget, a Swiss developmental psychologist, asserted that it is in human nature to be curious, inquisitive, and interested in novelty as they develop and acquire knowledge about their environment \cite{piaget}.
+Through observing human behaviour, it is clear that the ability to explore a domain is critical to the process of discovery and learning, regardless of the task being considered.
 
@@
 \end{equation}
-The base A2C algorithm does not explicitly seek random exploration as it learns to decide on the action at a https://arxiv.org/pdf/1601.06733.pdf state purely from the policy update.
-The most commonly used method to introduce randomness to the agent is through maximizing the entropy of the next-action sampling distribution.
+The base A2C algorithm does not explicitly seek random exploration as it learns to decide on the action at a state purely from the policy update.
+The most commonly used method to introduce randomness to the agent is through maximizing the entropy $E$, shown in Fig.
+\ref{eq:entropy}, of the next-action sampling distribution.
 This distribution is categorical with probabilities defined by the policy network logits.
 \begin{equation}\label{eq:entropy}
-\mathcal{L}_E = \sum_a (-\log \pi_a \times \mu_\pi)
+E = \sum_a (-\log \pi_a \times \mu_\pi)
 \end{equation}
@@
 \begin{equation}
-\mathcal{L} = \mathcal{L}_P + \lambda_{V}\mathcal{L}_V - \lambda_{E}\mathcal{L}_E + \mathcal{L}_{Fwd} + \mathcal{L}_{Inv}
+\mathcal{L} = \mathcal{L}_P + \lambda_{V}\mathcal{L}_V - \lambda_{E}E + \mathcal{L}_{Fwd} + \mathcal{L}_{Inv}
 \end{equation}
@@
 \end{subfigure}
-\caption{Here we compare the three A2C variants in the \texttt{Breakout} (left) and \texttt{Seaquest} (right) environments. Rewards are again cumulative over an entire episode.}
+\caption{Here we compare the three A2C variants in the \texttt{Breakout} (left) and \texttt{Seaquest} (right) environments.
+Rewards are again cumulative over an entire episode.
+We note that the performance of the curious learner (cyan) is worse at times than the A2C + Entropy model (yellow), which is perhaps due to the lack of hyper-parameter tuning and the density of the extrinsic reward in these environments.}
 \label{fig:seaquest-breakout-reward}
@@
 Specifically, \texttt{Seaquest} is most dense in that positive rewards are given for each ``kill'', which occurs very frequently due to the sheer number of enemies and the fact that the player's submarine is always firing projectiles at high velocity.
-In fact, positive rewards are already obtained in the early steps of training (as can be seen in Fig.
+In fact, positive rewards are already obtained in the early steps of training (as can be seen in the right of Fig.
 \ref{fig:seaquest-breakout-reward}).
@@
 
+\newpage
 \bibliography{main}

### 2020-04-17 plot.py
Update figs

@@ -36,7 +36,8 @@ for d in data:
     plt.plot(d[0], smooth(d[1], 0.95), color=colours[i], label=labels[i])
     i = (i+1) % 3
 
-# plt.ylim(0, 1800)
+plt.xlim(0, 1100000)
+plt.ylim(0, 40)
 plt.legend()
 plt.xlabel('Training Update Steps')
 plt.ylabel('Extrinsic Reward per Episode')
